---
title: "DATA 607  IS 607 – Project 2 1.Rmd"
author: "Yohannes Deboch"
date: "March 3, 2019"
output: html_document
---
```{r}

```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r}
library(tidyr)
library(dplyr)
library(ggplot2)
library(stringr)
```
Data 1
UNICEF under-five child mortality 
https://sejdemyr.github.io/r-tutorials/basics/data/unicef-u5mr.csv

 Create a .CSV file (or optionally, a MySQL database!) that includes all of the information
included in the dataset. You’re encouraged to use a “wide” structure similar to how the
information appears in the discussion item, so that you can practice tidying and
transformations as described below.
  Read the information from your .CSV file into R, and use tidyr and dplyr as needed to
tidy and transform your data. [Most of your grade will be based on this step!]

```{r}
dt <- read.csv("unicef-u5mr.csv")
```

Few first few rows
```{r}
head(dt)
```
View column names
```{r}
names(dt)
```
View data types
```{r}
sapply(dt,class)
```
Transform wide to long format
```{r}
dt_tidy <- gather(dt, Year,Value,-CountryName)
head(dt_tidy)
```
Extract Year and convert to numeric
```{r}
dt_tidy$Year <- str_extract(dt_tidy$Year,"\\d+$")
dt_tidy$Year <- as.numeric(dt_tidy$Year)
```



 Perform the analysis requested in the discussion item.

Average Child Mortality over the years
```{r}
dt_tidy %>%
    group_by(Year)%>%
    summarise(avg = mean(Value, na.rm = TRUE)) %>%
    ggplot(aes(Year,avg))+geom_line()+theme_classic()+ggtitle("AVerage Child Mortality Over The Years")
```
From the graph we can see that average child mortality has decreased over the years.

Top 5 country with highest child mortality
```{r}
dt_tidy %>%
    filter(Year == max(Year))%>%
    arrange(-Value)%>%
    head()
```
Angola,Chad,Somalia,Central African Republic ,Sierra Leone, Mali are the top country with highest child mortality in year 2015.
Top 5 country with lowest child mortality
```{r}
dt_tidy %>%
    filter(Year == max(Year))%>%
    arrange(Value)%>%
    head()
```
Luxembourg ,  Iceland,Finland ,  Norway , Slovenia, Cyprus are the top country with lowest child mortality in year 2015.

Conclusion: From the analysis we can say that child mortality has decrease gradually over the years. African countries has the highest child mortality and developed countries lowest child mortality.
Your code should be in an R Markdown file, posted to rpubs.com, and should include
narrative descriptions of your data cleanup work, analysis, and conclusions.