---
title: "DATA 607  IS 607 – Project 2 1.Rmd"
author: "Yohannes Deboch"
date: "March 3, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r}
library(tidyr)
library(dplyr)
library(ggplot2)
library(stringr)
```
Data 3
Population, total
https://data.worldbank.org/indicator/SP.POP.TOTL

 Create a .CSV file (or optionally, a MySQL database!) that includes all of the information
included in the dataset. You’re encouraged to use a “wide” structure similar to how the
information appears in the discussion item, so that you can practice tidying and
transformations as described below.
 Read the information from your .CSV file into R, and use tidyr and dplyr as needed to
tidy and transform your data. [Most of your grade will be based on this step!]

```{r}
pop <- read.csv("API_SP.POP.TOTL_DS2_en_csv_v2_10473719.csv", skip = 4)
```
View first few rows of the data
```{r}
head(pop)
```
Drop unnecessary variables
```{r}
pop <- pop %>%
    select("Country.Name",starts_with("X"),-X)
head(pop)
```

Remove unncessary variable X2018
```{r}
pop <- pop %>%
    select(-X2018)
head(pop)
```
Drop unnecessary rows

Convert wide to long format
```{r}
pop <- gather(pop, "Year","Population",-Country.Name)
head(pop)
```
Remove X from Year
```{r}
pop$Year <- str_replace(pop$Year,"X","")
```
See data type of the variables
```{r}
sapply(pop, class)
```
Correct data type by converting Year to Numeric
```{r}
pop$Year <- as.numeric(pop$Year)
```

See again data type of the variables 
```{r}
sapply(pop, class)
```
Rename variable names
```{r}
colnames(pop)[1] <- "Country"
```

Show first 10 rows
```{r}
head(pop,10)
```

 Perform the analysis requested in the discussion item.
Worlds total population change over the years
```{r}
pop %>%
    group_by(Year)%>%
    summarise("Total Population" = sum(Population, na.rm = TRUE))%>%
    ggplot(aes(Year,`Total Population`))+geom_line()+theme_classic()+ggtitle("World Population Change Over The years")+ylab("Population")
```

From the graph we can see that world populatin has increased over the years.

Top 10 countries with highest population in 2017
```{r}
pop %>%
    filter(Year == max(Year))%>%
    arrange(-Population)%>%
    head(30)
```


Top 10 countries with lowest population in 2017
```{r}
pop %>%
    filter(Year == max(Year))%>%
    arrange(Population)%>%
    head(10)
```
Conclusion:
We conclude that world population has increased over the years. China and India has the highest population in 2017.
Tuvalu has the lowest population in 2017

 Your code should be in an R Markdown file, posted to rpubs.com, and should include
narrative descriptions of your data cleanup work, analysis, and conclusions.